### 2023-05-01 chapter2/chapter2.tex
Moved outline into discussion and added more subsections to group related paragraphs

% CONCLUSION
% I. Summary of results
%    A. In this study, we demonstrate the first evidence of evolutionary constraint in IDRs across the \textit{Drosophila} proteome.
%    B. Most immediately, this work generates hypotheses of function related to sequence features that will permit the functional dissection of IDRs.
%    C. More broadly, however, it illustrates constraint of distributed features characterizes the evolution of some IDRs across multiple systems.

% II. Exchangeability and rate matrices
%    A. In addition to these global patterns of conserved features, we find IDRs exhibit other characteristic patterns of evolution.
%    B. For example, a comparison of the exchangeability matrices fit to the disorder and order regions show decreased exchangeability coefficients between
%    A. Interestingly, the exchangeability matrices suggest that although there is net flux towards disorder by depleting structure-promoting residues, disordered residues are more biochemically distinct in disordered regions than in structured regions. This suggests that although disorder-promoting amino acids are often grouped into a single category and described as largely hydrophilic, they have distinct roles in disordered regions. There is evidence that highly similar residues such as glutamine and asparagine have distinct properties in disordered regions that yielded different biophysical characters of phase-separated condensates, so this aligns with other observations. Conversely, structure promoting residues are more interchangeable. One hypothesis is as these bulkier hydrophobic residues are not constrained by packing, they are free to act more generically as small hydrophobic patches that mediate interactions in SLiMs and other binding regions in IDRs. Another model for this behavior is the stickers and spacers model of phase separation.
%    A. A caveat of this analysis is its dependence on sequence alignments which were created using scoring matrices that are in turn often derived from substitution models. Some approaches attempt to minimize the impact of this circular dependency through an EM-like procedure where substitution models are fit and the sequences are re-aligned with scoring matrices derived from those models in alternating rounds until convergence. Because our analyses fit the substitution models directly from alignments, the observed patterns we observe could be artifacts of using a scoring matrix largely fit to folded domains to align IDRs
%    A. However, as the exchangeability and rate matrices broadly reproduced the trends observed in previous studies which fit amino acid substitution models to disorder regions, these observations are likely not artifactual but instead represent reproducible differences in the substitution rates between residues.
%    B. Interpretations of exchangeability and rate matrices
%        1. Disorder-enriched residues are less interchangeable and play specific roles in these regions
%        2. Disorder-depeleted residues not constrained by packing and can act more generically as small hydrophobic patches
%            i. Stickers and spacers from Holehouse
%            ii. Staller papers and other high throughput mutational scanning of ADs
%            iii. Halfmann glutamine vs asparagine paper
%        3. Note about variances of parameters as shown in CV graphs

\section{Discussion}
Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Thus, the analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. We found that disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.} fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of intrinsic disorder.

@@

As discussed by Zarin \textit{et al.}, the interpretation of evolutionary signatures is complicated by several methodological caveats~\cite{Zarin2019}. For example, because IDRs are identified as contiguous segments of high predicted disorder, their boundaries are defined by adjacent structural elements. This approach can therefore split an IDR that is a single evolutionary or functional unit if it contains a semi-disordered module that scores below the threshold. Conversely, it can also merge two distinct IDRs if they are not separated by at least one folded domain. Another challenge is the significant overlap in the definitions of many features induces strong correlations that preclude straightforward quantitative manipulations or interpretations of an IDR's evolutionary signature. The original authors have since addressed this limitation in subsequent studies by applying machine learning methods to perform feature selection or learn features directly from alignments of IDRs~\cite{Zarin2021, Lu2022}. However, integrating these methods into a unified phylogenetic comparative framework will require further effort.

>>>>

\section{Discussion}
% I. Summary of results
%    A. In this study, we applied several phylogenetic models to IDRs to interrogate the evolution of their sequences and molecular features.
%    B. Most significantly, we demonstrate evidence of widespread evolutionary constraint in IDRs within the \textit{Drosophila} proteome
%    C. Furthermore, many IDRs can be grouped into discrete clusters on the basis of these evolutionary signatures
%    B. Most immediately, this work generates hypotheses of function related to sequence features that will permit the functional dissection of IDRs
%    C. More broadly, however, it illustrates constraint of distributed features characterizes the evolution of some IDRs across multiple systems.
%    D. However, these clusters require careful interpretation. The cluster descriptions summarize only a subset of the annotations with statistically significant enrichments. All clusters had many enriched annotations, but the summaries correspond to sets of enrichments with consisten . Because the significance tests were not corrected for multiple testing or controlled for their false discovery rate, many annotations are ,

% II. IDRs have characteristic patterns of amino acid substitution
%    A. In addition to these global patterns of conserved features, we find IDRs exhibit other characteristic patterns of evolution.
%    B. For example, a comparison of the exchangeability matrices fit to the disorder and order regions show decreased exchangeability coefficients between
%    A. Interestingly, the exchangeability matrices suggest that although there is net flux towards disorder by depleting structure-promoting residues, disordered residues are more biochemically distinct in disordered regions than in structured regions. This suggests that although disorder-promoting amino acids are often grouped into a single category and described as largely hydrophilic, they have distinct roles in disordered regions. There is evidence that highly similar residues such as glutamine and asparagine have distinct properties in disordered regions that yielded different biophysical characters of phase-separated condensates, so this aligns with other observations. Conversely, structure promoting residues are more interchangeable. One hypothesis is as these bulkier hydrophobic residues are not constrained by packing, they are free to act more generically as small hydrophobic patches that mediate interactions in SLiMs and other binding regions in IDRs. Another model for this behavior is the stickers and spacers model of phase separation.
%    A. A caveat of this analysis is its dependence on sequence alignments which were created using scoring matrices that are in turn often derived from substitution models. Some approaches attempt to minimize the impact of this circular dependency through an EM-like procedure where substitution models are fit and the sequences are re-aligned with scoring matrices derived from those models in alternating rounds until convergence. Because our analyses fit the substitution models directly from alignments, the observed patterns we observe could be artifacts of using a scoring matrix largely fit to folded domains to align IDRs
%    A. However, as the exchangeability and rate matrices broadly reproduced the trends observed in previous studies which fit amino acid substitution models to disorder regions, these observations are likely not artifactual but instead represent reproducible differences in the substitution rates between residues.
%    B. Interpretations of exchangeability and rate matrices
%        1. Disorder-enriched residues are less interchangeable and play specific roles in these regions
%        2. Disorder-depeleted residues not constrained by packing and can act more generically as small hydrophobic patches
%            i. Stickers and spacers from Holehouse
%            ii. Staller papers and other high throughput mutational scanning of ADs
%            iii. Halfmann glutamine vs asparagine paper
%        3. Note about variances of parameters as shown in CV graphs

\subsection{Disorder is correlated with many molecular features}
Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Thus, the analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. We found that disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.} fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of intrinsic disorder.

@@

\subsection{A workflow for future evolutionary analyses of IDRs}
As discussed by Zarin \textit{et al.}, the interpretation of evolutionary signatures is complicated by several methodological limitations~\cite{Zarin2019}. For example, because IDRs are identified as contiguous segments of high predicted disorder, their boundaries are defined by adjacent structural elements. This approach can therefore split an IDR that is a single evolutionary or functional unit if it contains a semi-disordered module that scores below the threshold. Conversely, it can also merge two distinct IDRs if they are not separated by at least one folded domain. Another challenge is the significant overlap in the definitions of many features induces strong correlations that preclude straightforward quantitative manipulations or interpretations of an IDR's evolutionary signature. The original authors have since addressed this in subsequent studies by applying machine learning methods to perform feature selection or learn features directly from alignments of IDRs~\cite{Zarin2021, Lu2022}. However, integrating these methods into a unified phylogenetic comparative framework will require further effort.
